Comparing variations in sediment transport capacity between subglacial and subaerial systems

### 2023-03-15 theory.tex
more small changes

\paragraph{Key findings:}
\begin{enumerate}
\item Variations in water discharge in subglacial systems are accommodated by water velocity of water, given the slow evolution of channel size.

@@

  To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
  Here, it is assumed that there is conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
  The shear stress $\tau_f$ at the river bed is represented as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW_tau}
      \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
    \end{equation}
  \end{linenomath*}
  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial systems.
  Channel width $w_{cf}$ is 
  \begin{equation}
    \label{eq:wcf}
    w_{cf} = k \, Q_w^e,
  \end{equation}
  % 
  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}.
  Cross-sectional area, $ S_f$, is 
  \begin{equation}
    \label{eq:Sf}
    S_f = \frac{Q_w}{v_f},
  \end{equation}
  % 
  where $v_f$ is water velocity, given by rearranging Equation \ref{eq:tau} as
  \begin{equation}
    \label{eq:vf}
    v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
  \end{equation}
  % 
  As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{multi_run.pdf}
      \caption{Ratio of range of a variable $\Gamma$ for glacial and subaerial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (subaerial) system. Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination. 
        Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
      \label{fig:range}
    \end{figure}
  \end{center}

@@

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to subaerial ones.
  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates the reduced shear stress in the subaerial system.
  No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.

  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S1).
  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure S1).

>>>>

\paragraph{Key findings:}
\begin{enumerate}
\item Water discharge variations in subglacial systems are accommodated by water velocity, not channel size, given the slow evolution of channels.

@@

  To parameterize the shear stress of water flowing across sediments in the subaerial channel,  the hydraulics parameterization presented in \citet{tucker1997} is implemented.
  Here, it is assumed that there is a conservation of mass and a sufficiently wide channel so that the hydraulic radius is consistent with flow depth, uniform flow, and the Darcy-Weisbach relationship.
  The shear stress $\tau_f$ at the river bed is represented as
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW_tau}
      \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q_w}{w_{cf}} \Big)^{\frac{2}{3}} \,\frac{\partial z_c}{\partial x}^{\frac{2}{3}},
    \end{equation}
  \end{linenomath*}
  where $\frac{\partial z_c}{\partial x}$ is the channel slope, and $f_p$ is the friction factor for subaerial systems.
  Channel width $w_{cf}$ is 
  \begin{equation}
    \label{eq:wcf}
    w_{cf} = k \, Q_w^e,
  \end{equation}
  % 
  where $k$ is a constant and $e$ is an exponent commonly equal to $\frac{1}{2}$ \citep{leopold1953}.
  Cross-sectional area, $ S_f$, is 
  \begin{equation}
    \label{eq:Sf}
    S_f = \frac{Q_w}{v_f},
  \end{equation}
  % 
  where $v_f$ is water velocity, given by rearranging Equation \ref{eq:tau} as
  \begin{equation}
    \label{eq:vf}
    v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
  \end{equation}
  % 
  As in Section~\ref{sect:sub_mode}, the width integrated shear stress is $\tau_{tf}=w_{cf}\,\tau_f$.
  Hydraulic diameter of the channel is given by $D_{hf} = \frac{4\,w_{cf}\,d}{2\,d+w_{cf}}$, with flow depth $d$ determined from knowledge of $S_f$ and $w_{cf}$ in a rectangular channel.

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{multi_run.pdf}
      \caption{Ratio of range of water velocity, shear-stress, and width integrated shear-stress for glacial and subaerial systems in response to mean water discharge (a, b, c), and  variability in water discharge (d, e, f).  
        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater than $1$ shown in white space are parameter combinations that vary more in a subglacial  system.
        Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
      \label{fig:range}
    \end{figure}
  \end{center}

@@

  The minimum ratio of  water velocity, $\frac{\chi(v_{g})}{\chi(v_{f})}$, and shear-stress, $\frac{\chi(\tau_{g})}{\chi(\tau_{f})}$, is around $2$, showing that these factors are substantially more variable in subglacial channels compared to subaerial ones (Figure~\ref{fig:range} g, h).
  The width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, spans between $1.02$ and $2.4$ because changing channel width accommodates some of the reduced shear stress variability in subaerial systems (Figure~\ref{fig:range} i).
  No runs experienced width integrated shear stress, $\frac{\chi(\tau_{tg})}{\chi( \tau_{tf})}$, less than $1$.

  Runs with the smallest variability in discharge ($\chi(Q_w)$) have the greatest variability in water velocity, shear stress, and width-integrated shear stress in subglacial systems compared to subaerial ones (Figure~\ref{fig:range} and S1).
  Yet, the greatest difference in the two systems at the smallest variations of water discharge still results in greater variability in the subglacial system compared to the subaerial one (Figure \ref{fig:range}).
